Interpretable Machine Learning

### 2019-01-15 chunks/story-wines.Rmd
continues slides

- ~6500 red and white Portuguese "Vinho Verde" wines
- Features: physiochemical properties
- Quality assessed by blind tasting, from 0 (very bad) to 10 (excellent)

@@

=> The random forest is the best model.

@@

## {data-background=../images/done-here.gif data-background-size=contain}
```{r, echo=FALSE, out.width='80%', include = FALSE}
knitr::include_graphics("../images/done-here.gif" )
```

@@

## What are the most important features?

## What are the most important features?

@@

## What are the most important features?

@@

```{r}
knitr::include_graphics("../images/pdp-rent.png")
```s

## How do features affect predictions?
```{r}
eff = FeatureEffect$new(pred, "alcohol", method = 'ale')
plot(eff)
```

## How do features affect predictions?
```{r}
eff = FeatureEffect$new(pred, "volatile.acidity", method = 'ale')
plot(eff)
```

@@

```{r}
inter = FeatureEffect$new(pred, feature = c("alcohol", "volatile.acidity"))
plot(inter)
```

@@

```{r}
shap = Shapley$new(pred, x.interest = worst)
plot(shap)
```

@@

```{r}
worst

worst2 = function(){
  w = worst
  w$volatile.acidity = 0.2
  w
}

>>>>

- ~6500 red and white Portuguese "Vinho Verde" wines
- Features: Physiochemical properties
- Quality assessed by blind tasting, from 0 (very bad) to 10 (excellent)

@@

=> Random forest has lowest Mean Absolute Error

@@

## {data-background=../images/dog-car.gif data-background-size=contain}

@@

## Which features are important?

## Which features are important?

@@

## Which features are important?

@@

## How do features affect predictions?
```{r}
knitr::include_graphics("../images/ale-intuition-1.jpg")
```

## How do features affect predictions?
```{r}
knitr::include_graphics("../images/ale-intuition-2.jpg")
```

## How do features affect predictions?
```{r}
knitr::include_graphics("../images/ale-intuition-3.jpg")
```

## How do features affect predictions?
```{r}
knitr::include_graphics("../images/ale-intuition-4.jpg")
```

## How do features affect predictions?
```{r}
knitr::include_graphics("../images/ale-intuition-5.jpg")
```

## Effect of Alchohol
```{r}
eff = FeatureEffect$new(pred, "alcohol", method = 'ale')
plot(eff)
```

## Effect of Volatile Aciditys
```{r}
eff = FeatureEffect$new(pred, "volatile.acidity", method = 'ale')
plot(eff)
```

@@

```{r}
inter = FeatureEffect$new(pred, feature = c("alcohol", "volatile.acidity"))
plot(inter, show.data = TRUE)
```

## Rule of thumb for wine quality?

@@

```{r}
shap = Shapley$new(pred, x.interest = worst)
plot(shap) + 
  scale_y_discrete("") + 
  scale_x_continuous("Feature contribution")
```

@@

```{r}
worst2 = function(){
  w = worst
  w$volatile.acidity = 0.2
  w
}

### 2019-01-15 chunks/tools.Rmd
continues slides

<div class="notes">
- Intrinsically interpretable models
- Tools for analyzing specific black box models
- Model-agnostic tools: analysing any black box model
- Interpretability through making data explainable
</div>

##

TODO: Overview (drawing)

- global vs. local
- interpretable model vs. post-hoc
- model-specific vs. model-agnostic
- data type image vs. text vs. tabular

@@

@@

TODO: Example for CNNs

TODO: Example for text (RNNs and attention?)

<div class="notes">
- Visualizing activations of neural networks
- Random Forest specific feature importance
- xgboost explainer
- problem: new models, not comparable between models, missing flexibility
</div>

@@

TODO: Drawing of feature effect

TODO: Drawing of importance

@@

TODO: Graphic for prototypes

>>>>

@@

## Model-specific Methods
Layerwise Relevance Propagation (LRP)

```{r, echo=FALSE}
knitr::include_graphics("../images/lrp.png")
```

<font size="2">
Bach, Sebastian, et al. "On pixel-wise explanations for non-linear classifier decisions by layer-wise relevance propagation." PloS one 10.7 (2015): e0130140.
</font>

@@

```{r, echo=FALSE}
knitr::include_graphics("../images/cnn-feature-viz.png")
```

<font size="2">
https://distill.pub/2017/feature-visualization/
</font>

@@

```{r}
knitr::include_graphics("../images/pdp-rent.png")
```

@@
